---
title: "Initial Korean covid 19"
author: "Ambar Arreola"
date: "6/21/2023"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata <- kRdata %>%
mutate(across(where(is.character), as.factor))

```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 2020
### This data was retrieved from a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
 
 
  
# Your Question - Is dying independent of sex?

## Methods

We will analyze the Sex of the patient and the result  if they are alive or dead.I will use the variables in any collection of cat~cat . Which is a type of problem that we can use the helper file for and we will copy into our project to get the answer if dying is independent of sex. 



## Graphical results - use ggplot commands to do this  



  #### Plot barplots ggplot way

```{r}
dd2 <- kRdata %>% group_by(result,Sex) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=result,y=count,fill=Sex))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=result,y=prcnt*100,fill=Sex)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

The data in this barplot allows us to see the results of the alive and dead people. We can see that Twice as many males died than females.The disease is killing more males than females according to the data from Korea.

  
## Numerical results

```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```

Of all the people that are still alive 56 are females and 36 are males. The number of dead people were 44 females and 64 males. The percentage was 98.7% of females remained alive and only 1.4% females died and 97% of males remained alive while 3% are dead. 

## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

In the group of 1831 were alive , in all more females are alive than males. The p value is 1 out of a thousand.It is a small chance to get a male and female dying.In the fisher test, 9 out of 10,000, the p value is about the same. You can also see the odds of male and females dying. Women have a 2.3% chance of surviving Covid-19 than men in Korea.

## Conclusions

In today's lesson we used the cat~cat file to create a plot using  gg plot way. We changed the class for result and named the project kRdata. We were also able to create a numerical graph and in percentage and counting. We also ran a chi-square test and a fisher's test and both were equal in p value. In the final results we were able to see the the odds of women surviving over men. The overall all results showed us that males had the short end of the stick because covid killed more males than females in Korea.
